Modeling costs and benefits of genomic wastewater surveillance

### 2023-05-15 paper/manuscript.qmd
updating manuscript

---
title: Modeling costs and benefits of genomic wastewater surveillance
# subtitle: Supplementary appendix draft
#author:
#  - name: Pedro Nascimento de Lima
#    email: plima@rand.org
#    affiliations: 
#        - id: rand
#          name: RAND Corporation
#          department: Engineering and Applied Sciences
#          address: 1776 Main Street
#          city: Santa Monica
#          state: CA
#          postal-code: 90401
#    attributes:
#        corresponding: true
#  - name: and colleagues
#    email: others@rand.org
#    affiliations:
#        - ref: rand
abstract: |
  Introduces a model to assess the value provided by wastewater genomic surveillance network. This document is meant to be used as a supplementary appendix of the paper.
keywords: 
  - Genomic Surveillance
  - Pandemic Response
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  elsevier-html:
    keep-tex: true
    mermaid:
      theme: neutral
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  docx:
      cite-style: super
  html:
    keep-tex: true
    mermaid:
      theme: neutral
---

@@

@@

where $f(x)$ is a function that maps a percent reduction in potentially infectious contacts to a monetary cost. At the time of the intervention, decision-makers do not have precise information on either the effects of interventions (expressed in our model as a percent reduction $100 * \tau$ in infectious contacts) or its costs represented by function $f(x)$. Even if public health decision makers could estimate the instantaneous costs of nonpharmaceutical interventions in real-time, they would not be able to observe the full costs of interventions that do not manifest during the time-frame of interventions (i.e., the life-time effect of school closures). Nevertheless, any attempt to value the benefits of surveillance systems require the specification of the costs of nonpharmaceutical interventions.

@@

@@

>>>>

---
title: Modeling costs and benefits of genomic wastewater surveillance
# subtitle: Supplementary appendix draft
#author:
#  - name: Pedro Nascimento de Lima
#    email: plima@rand.org
#    affiliations: 
#        - id: rand
#          name: RAND Corporation
#          department: Engineering and Applied Sciences
#          address: 1776 Main Street
#          city: Santa Monica
#          state: CA
#          postal-code: 90401
#    attributes:
#        corresponding: true
#  - name: and colleagues
#    email: others@rand.org
#    affiliations:
#        - ref: rand
abstract: |
  Introduces a model to assess the value provided by wastewater genomic surveillance network. This document is meant to be used as a supplementary appendix of the paper.
keywords: 
  - Genomic Surveillance
  - Pandemic Response
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  elsevier-html:
    keep-tex: true
    mermaid:
      theme: neutral
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  html:
    keep-tex: true
    mermaid:
      theme: neutral
---

@@

<!--
### Discrete-time stochastic implementation

The model described above is implemented as a discrete-time stochastic model using the following approach. We simulate outflows for each disease state vector $\mathbf{X}$ using a stochastic  $\mathbf{X_t} = \mathbf{X_{t-1}} \sim Binomial(\mathbf{X}, 1-exp(\mathbf{p}))$ hence.

-->

@@

where $f(x)$ is a function that maps a percent reduction in potentially infectious contacts to a monetary cost. $f(x)$ is monotonically increasing with $\mathbf{l}_t\tau$, and may exhibity decreasing marginal returns to a reduction in transmissibility. At the time of the intervention, decision-makers do not have precise information on either the effects of interventions (expressed in our model as a percent reduction $100 * \tau$ in infectious contacts) or its costs represented by function $f(x)$. Even if public health decision makers could estimate the instantaneous costs of nonpharmaceutical interventions in real-time, they would not be able to observe the full costs of interventions that do not manifest during the time-frame of interventions (i.e., the life-time effect of school closures). Nevertheless, any attempt to value the benefits of surveillance systems requires the specification of the costs of nonpharmaceutical interventions.

@@

<!--

@@

-->
